---
title: "Joint Type A and Type B Testing (Wolak, 1989)"
author:
  - "Leonard Vanbrabant"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Joint Type A and Type B Testing (Wolak, 1989)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  tidy = FALSE
)
library(restriktor)
```

# When do you need a joint test?

A common research claim with order restrictions is: *"there is demonstrable
development in the predicted direction"*. With `iht()` this claim rests on
**two** hypothesis tests for the same set of restrictions
$R\theta \ge rhs$:

- **Type A** — H0: $R\theta = rhs$ vs. HA: $R\theta \ge rhs$. Rejecting H0
  shows there is *some* effect in the restricted direction. The test
  statistic is $T_A = (\tilde{\theta} - \bar{\theta})'
  \Sigma^{-1} (\tilde{\theta} - \bar{\theta})$ (EI), the distance between
  the equality-restricted and the order-restricted estimate.
- **Type B** — H0: $R\theta \ge rhs$ vs. an unrestricted alternative.
  *Not* rejecting H0 shows the data do not contradict the predicted
  direction. The test statistic is $T_B = (\hat{\theta} - \tilde{\theta})'
  \Sigma^{-1} (\hat{\theta} - \tilde{\theta})$ (IU), the distance between
  the unrestricted and the order-restricted estimate.

The substantive conclusion "development in the predicted direction" requires
type A to reject *and* type B not to reject — a statement about the **pair**
of outcomes. Reporting two separate p-values does not quantify how likely
that pair is under the shared null hypothesis, and multiplying the two
marginal levels ($\alpha_A \cdot \alpha_B$) is wrong: the two statistics are
computed from the same data and are *dependent*.

Wolak (1989, Theorem 1) derived the joint distribution of $(T_B, T_A)$ under
the least favorable null configuration $R\theta = rhs$, which both
hypotheses share:

$$
\Pr[T_B \ge c_1,\; T_A \ge c_2] \;=\; \sum_{k=0}^{L} w(L, k, \Omega)\,
\Pr[\chi^2_{q-k} \ge c_1]\, \Pr[\chi^2_k \ge c_2],
$$

where $q$ is the total number of constraints, $L$ the number of inequality
constraints, and $w(L, k, \Omega)$ the chi-bar-square weights that
restriktor already uses for the marginal p-values. Conditional on $k$ the
two statistics are independent, but marginally they are not. For linear
models with an estimated error variance, both statistics share the same
variance estimate and the exact finite-sample version replaces the
chi-square tails by an F-type mixture (Wolak, 1987); `iht_joint()` uses
that version, consistent with the marginal p-values of `iht()`.

# Example: ordered group means

The `ZelazoKolb1972` data contain ages (in months) at which infants started
to walk. The hypothesis is that walking exercises do not *increase* the mean
age: active exercises should give the lowest mean age, no exercises the
highest.

```{r}
DATA <- subset(ZelazoKolb1972, Group != "Control")
fit.lm <- lm(Age ~ -1 + Group, data = DATA)
coef(fit.lm)
```

The predicted order is a set of sign restrictions on differences:

```{r}
myConstraints <- ' GroupActive < GroupPassive < GroupNo '
out <- iht_joint(fit.lm, constraints = myConstraints)
out
```

The output shows the two marginal tests (identical to
`iht(..., type = "A")` and `iht(..., type = "B")`), the joint tail
probability $\Pr[T_B \ge t_B, T_A \ge t_A]$ under the shared null, and the
combined decision at the chosen `alpha`:

- region **(a)**: type A does not reject — H0: $R\theta = rhs$ is retained;
- region **(b)**: type A rejects, type B does not — support for
  $R\theta \ge rhs$ with $R\theta \ne rhs$: "development in the predicted
  direction";
- region **(c)**: type B rejects — the order restrictions themselves are
  rejected.

The joint tail probability is the single number that expresses how
exceptional the observed *pair* of statistics is if in truth
$R\theta = rhs$: it replaces the informal combination of two separate
p-values with one statistically correct measure.

# Why the product of two p-values is wrong

The dependence between $T_A$ and $T_B$ makes the naive product of marginal
levels systematically wrong. A small numerical demonstration with the
components of the Wolak formula:

```{r}
fit.con <- restriktor(fit.lm, constraints = myConstraints)
wt <- iht_joint(fit.con)$wt.bar    # chi-bar-square weights
q  <- nrow(fit.con$constraints)    # number of constraints
df2 <- fit.con$df.residual

# marginal 5% critical values for both tests: setting one argument of
# pfbar_joint() to 0 gives the marginal tail probability of the other test
cA <- uniroot(function(c) {
  pfbar_joint(0, c, df_total = q, df2 = df2, wt.bar = wt) - 0.05
}, c(1e-8, 500))$root
cB <- uniroot(function(c) {
  pfbar_joint(c, 0, df_total = q, df2 = df2, wt.bar = wt) - 0.05
}, c(1e-8, 500))$root

# joint probability of exceeding BOTH critical values under Rb = r:
pfbar_joint(cB, cA, df_total = q, df2 = df2, wt.bar = wt)

# the naive product suggests:
0.05 * 0.05
```

The exact joint probability differs clearly from $0.05^2 = 0.0025$. The
reason is visible in the Wolak formula: conditional on $k$, large values of
$T_A$ (which favor $\chi^2_k$ components with large $k$) go together with
*small* values of $T_B$ (which then has few degrees of freedom left,
$\chi^2_{q-k}$), so the two statistics are negatively related under the
null. Treating them as independent misstates the joint error rate — which is
exactly the rate that matters for the combined claim "type A rejects and
type B does not".

# References

Kudo, A. (1963). A multivariate analogue of the one-sided test.
*Biometrika*, 50, 403--418.

Silvapulle, M.J., and Sen, P.K. (2005). *Constrained Statistical
Inference*. Wiley, New York.

Wolak, F. (1987). An exact test for multiple inequality and equality
constraints in the linear regression model. *Journal of the American
Statistical Association*, 82, 782--793.

Wolak, F. (1989). Local and global testing of linear and nonlinear
inequality constraints in nonlinear econometric models. *Journal of
Econometrics*, 41, 205--235.
